\papertitle

### 2014-03-17 final-report/paper.tex
section on graph

\documentclass{chi2009}
\usepackage{times}
\usepackage{url}
\usepackage{graphics}
\usepackage{color}
\usepackage[pdftex]{hyperref}
\usepackage{xspace}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}

@@

% What is myria, how does it work (look at report from last quarter but shorter), which languages are supported
% motivation
% questions (inspiration: https://docs.google.com/a/uw.edu/document/d/1Syn65bu_p6S4RQETaozsleYcRBauQJ7IduYrctS1wrw/edit?pli=1#)
% why is a visualization a good way to answer the questions

@@

We found that the performance of the query might be affected by the following
factors, which greatly influenced the design of our visualization tool:
\begin{itemize}
   \item \emph{Wrong/unoptimized physical query plan.} We manually analyzed
several physical query plans and discovered that some queries were poorly
optimized. This led us to design a view that allows interactive exploration of
the physical query plan (mapped to a graph of query operators). 
   \item \emph{Stragglers, tail latency, expensive query operators, poor
scheduling.} Our targeted systems are \emph{distributed} databases. Therefore,
as the query is executed in a distributed environment, the query coordinator
has to wait until all the workers finish executing their assigned tasks. To
give insight into how the tasks are distributed and executed at each workers,
we designed a view that shows the cluster utilization over time and at each
worker.
   \item \emph{Poor data partitioning, data skew.} Network traffic can cause
serious delay when executing a distributed query. Due to poor data
partitioning, workers might need to send unnecessary big amount of data between
each other. We designed a view that allows the developer/programmer to analyze
the data traffic generated while executing the query.  
\end{itemize}

@@

\system's architecture (Fig.~\ref{fig:arch}) consists of a back-end,
used as a plug-in interface for the distributed database system targeted, and a
front-end which implements the views described above. The front-end and
back-end of \system are highly decoupled so that the visualization can be used not only with
Myria but also with other systems such as Spark or Hadoop. In
Section~\ref{sec:back} we illustrate how we implemented log collections and
aggregation in Myria.

@@

\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{images/viz_arch}
  \caption{Overview of the log collection and transformation architecture. Raw event logs are collected on each worker. The data is stored in relations. To download the data, a query has to be executed.}
  \label{fig:arch}
\end{figure}

@@

\subsection{Front-end}

The Myria web front-end server is written in Python and runs on Google App Engine. \system's user-interface (UI) is embedded into Myria's web front-end. We build \system's UI using D3\cite{2011-d3} to deliver an interactive visualization experience to the user. D3 is a JavaScript framework for data visualziation on the web. We used visualization techniques such as \emph{focus+context}\cite{furnas1986generalized} that possible to implement conveniently using D3.

@@

\system's web UI is divided into two components: (1) the context window which provides a visualization of the \graph, and (2) the performance window which provides some insight into an element selected in the context view. The context window contains a view of the \graph, rendered as a graph. The user can navigate the \graph by expanding query fragments into the operators that compose it. The user can chose to select a fragment of interest which will render a \fragment visualization for the selected fragment in the performance window. Alternatively, the user can select an fragment-to-fragment edge in the \graph thus rendering a \network visualization for the selected edge in the performance window. If no elements are selected in the \graph view, an \overall visualization is rendered in the performance window, displaying the aggregate worker utilization over time for each fragment.

@@

@@

The \fragment view comprises two charts: the aggregation chart and the workers chart.

\begin{figure}[ht]
  \centering
  \begin{subfigure}[b]{0.49\columnwidth}
    \includegraphics[width=\columnwidth]{images/skew}
    \caption{Data skew.}
    \label{fig:skew}
  \end{subfigure}
  \begin{subfigure}[b]{0.49\columnwidth}
    \includegraphics[width=\columnwidth]{images/no-skew}
    \caption{No data skew.}
    \label{fig:no-skew}
  \end{subfigure}
  \caption{The network view shows a distinctive pattern when there is data skew. Figure~\ref{fig:skew} shows that most tuples are sent to worker 1 and 5 (origin: x-axis, destination: y-axis).}
  \label{fig:network}
\end{figure}

>>>>

\documentclass{chi2009}
\usepackage{times}
\usepackage{url}
\usepackage{graphics}
\usepackage{color}
\usepackage[pdftex]{hyperref}
\usepackage{xspace}

\usepackage{graphicx}

@@

% What is myria, how does it work (look at report from last quarter but shorter), which query languages are supported
% motivation
% questions & goals
% why is a visualization a good way to answer the questions
% contributions

@@

We found that the performance of the query might be affected by the following
factors, which influenced :
\begin{itemize}
   \item 
\end{itemize}

@@

\begin{itemize}
    \item determine possible causes (data skew, stragglers, etc.) for why a query might be slow and design visualizations to
help
    \item we designed three views - physical query plan, fragment view, communication view.
\end{itemize}

\system's frontend is embedded into Myria's web frontend and provides comprehensive performance views to the user. There are 4 views in total that allow the user to visualize query performance details under different angles:
\begin{itemize}
    \item Physical Query Plan view: this view allows the user to connect the query she wrote with the physical query plan that the query optimizer generates.
    \item etc. 
\end{itemize}

The front and back end of \system are highly decoupled so that it can be used not only with Myria but also with other systems such as Spark or Hadoop. In Section~\ref{sec:back} we illustrate how we implemented log collections and aggregation in Myria.

@@

\begin{figure}[ht]
  \includegraphics[width=\columnwidth]{viz_arch}
  \caption{Architecture}
  \label{fig:arch}
\end{figure}

@@

\subsection{Front-end}
The Myria web front-end server is written in Python and runs on Google App Engine. \system's user-interface (UI) is embedded into Myria's web front-end. We build \system's UI using D3 to delive an interactive visualization experience to the user. D3 is a JavaScript framework for data visualziation on the web. We used visualization techniques such as \emph{focus+context} that possible to implement conveniently using D3.

@@

\system's web UI is divided into two components: (1) the browser panel which provides a visualization of the \graph, and (2) the performance panel which provides some insight into an element selected in the browser view. The browser panel contains a view of the \graph, rendered as a graph. The user can navigate the \graph by expanding query fragments into the operators that compose it. The user can chose to select a fragment of interest which will render a \fragment visualization of the selected fragment in the performance panel. Alternatively, the user can select an fragment-to-fragment edge in the \graph thus rendering a \network visualization of the selected edge in the performance panel. If no elements are selected in the \graph view, an \overall visualization is rendered in the performance panel, displaying the aggregate worker utilization over time for each fragment.

@@

The \graph view provides context to the user. It provides a visualization of the physical query plan. A query plan is an ordered set of steps used to access data in database management systems. The physical query plan is represented by a graph where each node represents a query fragment, and each link represents inter-worker communication from one query fragment to the next. Each query fragment is a collection of query operators. One can think of a query fragment as a job unit that gets scheduled on a cluster of machines. Consequently transitioning from one query fragment to another generally incurs all-to-all communication between worker nodes on the cluster.

The user can perform two classes of actions that will render a new visualization in the performance panel:
\begin{itemize}

    \item \textbf{Empty selection:} by default if no fragments or edges are selected in the \graph view, an \overall visualization is rendered in the performance panel. This visualization displays the aggregate worker utilization over time for each fragment and provides cluster utilization data at a glance.
    \item \textbf{Fragment selection:} when selecting a fragment, a \fragment visualization gets displayed in the performance panel. The \fragment visualization provides aggregate cluster utilization information complemented by per-worker task schedules. When a fragment is selected, the operands inside of the fragment in the \graph view are color-coded to allow the user to easily match each task in the per-worker schedule in the performance window with the corresponding query operator.
    \item \textbf{Edge selection:} upon selecting one or more fragment-to-fragment edges, a \network visualization gets displayed, providing aggregate exchange data between cluster nodes.

\end{itemize}

\textbf{Graph Rendering:} We use D3 to render the \graph graph, which allows us to support various interactions and transistions. We use GraphViz \cite{graphviz} in the backend to generate graph layout information. GraphViz allows us to generate graph layouts optimized for space. The layout information is then fed into a D3-based rendering engine.

\textbf{Graph Navigation:} A query plan can be arbitrarily large. Thus we offer two mechanisms that facilitate exploration of the graph for the user: (1) expanding/reducing fragments and (2) zooming. The first mechanism can reduce the size of the graph by a constant factor by collapsing operators that compose a fragment into a single node. The user can click once on a collapsed fragment to expand it, and click once on an expanded fragment to collapse it. Expanding or reducing a node can cause large changes in the graph layout because GraphViz optimizes the graph layout to minimize area. We implemented transitions to allow the user to track the fragments as they get shuffled around after one fragment gets expanded or reduced. The second mechanism that facilitates exploration of the graph for the user is zooming. This feature was implemented in D3 and allows the user to quickly jump from one part of the graph to another, by zooming in and out of the view.

\textbf{Tooltip Information:} TODO

@@
